---
title: "Curling"
author: "Owen Raffa"
date: "2026-01-10"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("tidyverse")
library("tidymodels")
library("ggforce")
library("ggplot2")
```

```{r}
ends <- read.csv("Ends_Dummy.csv") %>%
  as_tibble()
ends
```

```{r}
model <- lm(Result ~ PowerPlayUsed, data = ends)
summary(model)
```

PowerPlayUsed Coefficient = 0.70886
Using a power play is associated with about +0.71 more points per end, 
on average, compared to not using one. Teams without a power play score about 
0.93 (intercept) points per end, while teams using a power play score about 
1.64 points per end.

```{r}

boxplot(Result ~ PowerPlayUsed,
        data = ends,
        names = c("No Power Play", "Power Play"),
        ylab = "Points Scored per End",
        main = "Power Play Advantage in Mixed Doubles Curling",
        col = c("#FF4D4D", "#4DA6FF"),
        border = "black")
```

The boxplots compare points scored per end when teams use a power play versus 
when they do not. The entire distribution shifts upward when a power play is 
used: the median score is higher, the middle 50% of outcomes are higher, and 
there are more high-scoring outliers. This shows that teams not only score more 
on average with a power play, but also have a greater chance of producing big scoring ends. In contrast, non-power-play ends are clustered around zero or one point. Together, these patterns visually demonstrate a clear comparative
advantage to using the power play.